### 2022-10-04 Paper.Rmd
Fixed some grammer/spelling errors

To create a great dataset, changes need to be made to the downloaded dataset. The dataset has been gathered of 1484 yeast sequences from SWISS-PROT using the annotations from YPD. In this section, we are going to talk about the results we found from the dataset.

@@

Below a table with a description of all the attributes abbreviation, explanation and data types.

@@

The last column is the sequence's localization site. There are 10 different possibilities for this. Yeast protein were classified into 10 classes: **cytoplasmic:** cytoskeletal (CYT); nuc]ear (NUC); vacuolar (VAC); mitochondrial (MIT); isomal (POX); **extracellular:** including those localized against the cell wall (EXC); proteins localized to the lumen of the endoplasmic reticulum (ERL); membrane proteins with a cleaved signal (ME1); membrane proteins with an uncleared signal (ME2); and membrane proteins with no N-terminal sign (ME3), where ME1, ME2,and ME3 proteins may be localized to the plasma membrane, the endoplasmic reticulum membrane, or the membrane of a golgi body.

Here are the 10 in question:

@@

There are alot of CYT and NUC localization. ERL localization is the least. There are only 5 of these in the dataset.

@@

8 features were used in classification: the presence or absence of an HDEL pattern as a signal for retention in the endoplasmic reticulum lumen (erl); the result of discriminant analysis on the amino acid content of vacuolar and extracellular proteins (vac); the result discriminant analysis on the amino acid composition of the 20-residue N-terminal region of mitochondrial and non-mitochondrial proteins (mit); the presence absence of nuclear localization consensus patterns combined with a term reflecting the frequency of basic residues (nuc); and some combination of the presence of a short sequence motif and the result of discriminant analysis of the amino acid composition of the protein sequence (pox).

@@

\begin{center}
\emph{Table 3:} First 6 rows of the dataset.
\end{center}

As seen from the table the loaded dataset. Has a char datatype for the first and last column. And every other column has a double as datatype.

@@

Figure 1 shows the amount of each clasification. There are alot of CYT and NUC localization. And there are only `r nrow(subset(data, loc.site=="ERL", loc.site))` of ERL.

@@

The first column has been dropped since it is not necessary. Since the sequence names contribute nothing to create a prediction model.

@@

\begin{center}
\emph{Table 4:} First 6 rows of the dataset with the first column dropped.
\end{center}

@@

The data type of the column loc.site has been succesfully changed to factors.

The ERL column has changed. From an double to a logical datatype. As seen in figure 2 below, only numbers 1 and 0.5 appear, but for a logical datatype you need 1 and 0. All 0.5 have been changed to 0.

@@

As observed in table 6 the data has already been transformed with a min-max normalisation. All the datapoints are between 0 and 1.Mcg and gvh are normally distributed. Alm and mit are skewed but nothing to crazy to be worried about. Vac and nuc are really skewed however, not yet bad enough to worry about it. Pox is very oddly distributed.

@@

Almost all the datapoint are 0. This could cause problems later on. Since more than 95% of all the datapoints are 0.

@@

If the pox value is not 0 then the classification is Pox, Mit or Cyt. This means that Pox has a high probability when the Pox value is not 0.

@@

This has been tranformed to a long matrix.

@@

As seen from the heatmap there is some correlation between mcg and gvh. While on the contrary Pox barely has any correlation with a other variable.

There needs to be data correlation with the classes. The heigth of the peak doesn’t matter much, shifted peaks do.

@@

There are shifted peaks. this shows a distribution of the different classes.

@@

As found in the heatmap Mcg and Gvh are clustered. Vac and Alm are somewhat clustered they are close to eachother. Nuc, Mit and Pox are all on their own.

@@

Mcg and Gvh are also clustered in this PCA plot. As in contrast with the MDS plot Vac and Alm are not clustered at all. Pox is in the centre.

@@

Our research question is: Can machine learning predict protein localisation based on their aminoacid sequences. And to answer this question properly, a clean dataset is needed to create a model. The results will show what is wrong with the data and the correlations between the different classifications. The results are not completely trustworthy because there is too little of them. Namely, there are only 1484 oberservations. This is too few for prediction with 10 different classifications.

@@

Nor was the Erl column considered. This indicates the presence of "HDEL" substring. If this had been looked at, more correlations would have emerged. Which could later become significant. There was also something wrong with the ERL column. As stated on the site, it is a binary/logical type. Either 0 or 1. It is present or it is not present. However, there were only values of 0.5 and 1. All values of 0.5 were changed to 0.

In the case of the Pox column, something strange was noticed. Out of 1484 datapoints, 15 are not zero, 11 of them have as classifications Pox. This means that if the value is not zero that the classification is highly probable pox.

@@

The aim was to create a clean dataset, to be used for our machine learning project. There is a clean dataset with no missing values and the correct data types. The data gives a clear overview of a protein classification and their attributes. However, not much was available. With only 1484 observations, this is just too little. And there is a very big difference between the number of observations per classifications. For ERL there are only 5 observations and for CYT there are 463. There is too big a difference between these. By improving this dataset over time, the prediction can be perfected even more. These constant improvements are important to find.

>>>>

To create a great dataset, changes need to be made to the downloaded dataset. The dataset consists of 1484 yeast sequences from SWISS-PROT using the annotations from YPD. In this section, we are going to talk about the results we found from the dataset.

@@

Below is a table with a description of all the attributes' abbreviations, explanations, and data types.

@@

The last column is the sequence's localization site. There are ten different possibilities for this. Yeast proteins were classified into ten classes: **cytoplasmic:** cytoskeletal (CYT); nuc]ear (NUC); vacuolar (VAC); mitochondrial (MIT); isomal (POX); **extracellular:** including those localized against the cell wall (EXC); proteins localized to the lumen of the endoplasmic reticulum (ERL); membrane proteins with a cleaved signal (ME1); membrane proteins with an uncleared signal (ME2); and membrane proteins with no N-terminal sign (ME3), where ME1, ME2,and ME3 proteins may be localized to the plasma membrane, the endoplasmic reticulum membrane, or the membrane of a golgi body.

Here are the ten in question:

@@

There are alot of CYT and NUC localizations. ERL localization is the least. There are only five of these in the dataset.

@@

Eight features were used in classification: the presence or absence of an HDEL pattern as a signal for retention in the endoplasmic reticulum lumen (erl); The results of discriminant analysis on the amino acid content of vacuolar and extracellular proteins (vac); the result of discriminant analysis on the amino acid composition of the 20-residue N-terminal region of mitochondrial and non-mitochondrial proteins (mit); the presence or absence of nuclear localization consensus patterns combined with a term reflecting the frequency of basic residues (nuc); and some combination of the presence of a short sequence motif and the result of discriminant analysis of the amino acid composition of the protein sequence (pox).

@@

\begin{center}
\emph{Table 3:} First six rows of the dataset.
\end{center}

As seen from the table, the loaded dataset. It has a char datatype for the first and last column. And every other column has a double datatype.

@@

Figure 1 shows the number of each classification of the dataset. There are alot of CYT and NUC localization. And there are only `r nrow(subset(data, loc.site=="ERL", loc.site))` of ERL.

@@

The first column has been dropped since it is not necessary. Since the sequence names contribute nothing to creating a prediction model.

@@

\begin{center}
\emph{Table 4:} First six rows of the dataset with the first column dropped.
\end{center}

@@

The data type of the column loc.site has been successfully changed to factors.

The ERL column has changed. from a double to a logical datatype. As seen in figure 2 below, only numbers 1 and 0.5 appear, but for a logical datatype you need 1 and 0. All 0.5 have been changed to 0.

@@

As observed in table 6 the data has already been transformed with a min-max normalisation. All the datapoints are between 0 and 1. Mcg and gvh are normally distributed. Alm and mit are skewed, but nothing too crazy to be worried about. Vac and nuc are really skewed. However, it's not yet bad enough to worry about it. Pox is very oddly distributed.

@@

Almost all the data points are 0. This could cause problems later on. Since more than 95% of all the data points are 0.

@@

If the pox value is not zero then the classification is Pox, Mit or Cyt. This means that Pox has a high probability when the Pox value is not zero.

@@

This has been transformed into a long matrix.

@@

As seen from the heatmap, there is some correlation between mcg and gvh. Pox, on the other hand, has almost no correlation with any other variable.

There needs to be data correlation with the classes. The height of the peak doesn’t matter much, shifted peaks do.

@@

There are shifted peaks. This shows the distribution of the different classes.

@@

As found in the heatmap, Mcg and Gvh are clustered. Vac and Alm are somewhat clustered; they are close to each other. Nuc, Mit and Pox are all on their own.

@@

Mcg and Gvh are also clustered in this PCA plot. As such, in contrast with the MDS plot, Vac and Alm are not clustered at all. Pox is in the centre.

@@

The research question is: can machine learning predict protein localization based on their aminoacid sequences? And to answer this question properly, a clean dataset is needed to create a model. The results will show what is wrong with the data and the correlations between the different classifications. The results are not completely trustworthy because there is too little of them. Namely, there are only 1484 oberservations. This is too few for prediction with 10 different classifications.

@@

Nor was the Erl column considered. This indicates the presence of the "HDEL" substring. If this had been looked at, more correlations would have emerged. Which could later become significant. There was also something wrong with the ERL column. As stated on the site, it is a binary/logical type. Either 0 or 1. It is present or it is not present. However, there were only values of 0.5 and 1. All values of 0.5 were changed to 0.

In the case of the Pox column, something strange was noticed. Out of 1484 datapoints, 15 are not zero. 11 of them have the classification of Pox. This means that if the value is not zero, the classification is highly probable Pox.

@@

The aim was to create a clean dataset to be used for our machine learning project. There is a clean dataset with no missing values and the correct data types. The data gives a clear overview of the protein classification and their attributes. However, not much was available. With only 1484 observations, this is just too little. And there is a very big difference between the number of observations per classification. For ERL there are only 5 observations, and for CYT there are 463. There is too big a difference between these. By improving this dataset over time, the prediction can be perfected even more. These constant improvements are important to find.
